---
title: "bayes-ch6"
output: html_document
---

# Chapter 6

### Exercise 6.1-A

```{r}
source("DBDA2Eprograms/DBDA2E-utilities.R")
source("DBDA2Eprograms/BernBeta.R")

posterior <- BernBeta(priorBetaAB = c(4,4), Data=c(1), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6.2-B

```{r}
posterior <- BernBeta(priorBetaAB = posterior, Data=c(1), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6.2-C

```{r}
posterior <- BernBeta(priorBetaAB = posterior, Data=c(0), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6.2-D

```{r}
posterior <- BernBeta(priorBetaAB = c(4,4), Data=c(0,1,1), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

The results are the same (between 6.2-C and 6.2-D).

### Exercise 6.2-A

In a survey, 58 would like candidate A (1) and 42 would like candidate B (0), given a uniform prior, what the HDI posterior?
```{r}
posterior <- BernBeta(priorBetaAB = c(1,1), Data=c(rep(1,58), rep(0,42)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

poll 100 new people, 57 would like candidate A (1), 43 for B (0).

```{r}
posterior <- BernBeta(priorBetaAB = posterior, 
                      Data=c(rep(1,57), rep(0,43)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6-3 
First: 40 people chose F and 10 chose J.  
Second: 15 chose F and 35 chose J.
assume uniform prior distribution

Set F = 1 and J = 0

```{r}
posterior <- BernBeta(priorBetaAB = c(1,1), 
                      Data=c(rep(1,40), rep(0,10)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
posterior <- BernBeta(priorBetaAB = posterior, 
                      Data=c(rep(1,15), rep(0,35)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6-4

```{r}
posterior <- BernBeta(priorBetaAB = c(0.1,0.1), 
                      Data=c(rep(1,4), rep(0,1)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6-5-A

```{r}
### Strong bias towards a fair coin
posterior <- BernBeta(priorBetaAB = c(100,100), 
                      Data=c(rep(1,9), rep(0,1)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

### Exercise 6.5-B

```{r}
### Strong bias towards a fair coin
posterior <- BernBeta(priorBetaAB = c(0.1,0.1), 
                      Data=c(rep(1,9), rep(0,1)), plotType="Bars",
                      showCentTend = "Mode", showHDI = TRUE, showpD = FALSE)
```

