---
title: "bayes-ch5"
output: html_document
---


```{r echo=F, warning=FALSE, message=FALSE} 
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(tidyr))
#suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(pbapply))
suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, fig.width=8, fig.height=6)
```

# Chapter 5

### Exercise 5.1

To solve exercise 5.1, let's first make a function to calculate the probabilities given a prior.

We assume:

* p(Test = pos | $\theta$ = sick) = 0.99 and
* p(Test = pos | $\theta$ = well) = 0.05  (false positve)

```{r}
getposterior <- function(prior) {
  basetable <- data.frame(sick=c(0.99 * prior, 0.01 * prior),
                       well=c(0.05 * (1-prior), 0.95 * (1-prior)))
  row.names(basetable)=c("pos", "neg")
  basetable
}
```

Now we can set the prior to 0.001  (1 in 1,000 people have disease) and then pull the table given that prior.

```{r}
basetable <- getposterior(0.001)
kable(basetable)
```

Now if we have one negative test result what is the probibility they have the disease?

* p($\theta$ = sick | Test = pos) 

```{r}
prior1 <- basetable["pos", "sick"] / (basetable["pos", "sick"] + basetable["pos", "well"])
prior1  # should be 0.19
```

Now what if we run a second test and the result is negative, what is the probability they are sick? 

* p($\theta$ = sick | Test = neg), prior = `r prior1` 

```{r}
basetable <- getposterior(prior1)
kable(basetable)
```

and...

```{r}
prior2 <- basetable["neg", "sick"] / (basetable["neg", "sick"] + basetable["neg", "well"])
prior2
```

### Exercise 5.2-A

Given population of 100,000, compute expected frequency of people in each cell

```{r}
kable(getposterior(0.001)*100000)
```

### Exercise 5.2-B

From frequencies, determine proportion of people who have the disease, given that their test result is positive.  Intuitively 99 / 4995+99, so about 100/5100 or 1/51 or 0.02.

```{r}
basetable <- getposterior(0.001)*100000
basetable["pos", "sick"] / (basetable["pos", "sick"] + basetable["pos", "well"])
```

### Exercise 5.2-C

```{r}
basetable <- getposterior(0.001)*10000000
# p(D = pos | sick)
basetable["pos", "sick"] / (basetable["pos", "sick"] + basetable["pos", "well"])
# p(D = neg | sick)
basetable["neg", "sick"] / (basetable["neg", "sick"] + basetable["neg", "well"])
# next tree
# p(D = pos | well)
basetable["pos", "well"] / (basetable["pos", "sick"] + basetable["pos", "well"])
# p(D = neg | well)
basetable["neg", "well"] / (basetable["neg", "sick"] + basetable["neg", "well"])
```

### Exercise 5.2-D

Given the above table, what proportion of people who test positive at first and then negative on retest actually have the disease?

```{r}
newprior <- basetable["pos", "sick"] / (basetable["pos", "sick"] + basetable["pos", "well"])
basetable <- getposterior(newprior)
basetable["neg", "sick"] / (basetable["neg", "sick"] + basetable["neg", "well"])
```

### Exercise 5.3-A

test is negative, what is the probability that person is sick?

```{r}
basetable <- getposterior(0.001)
basetable["neg", "sick"] / (basetable["neg", "sick"] + basetable["neg", "well"])
```

### Exercise 5.3-B

Then retested, and test is positive, what's the probability they are sick?

```{r}
newprior <- basetable["neg", "sick"] / (basetable["neg", "sick"] + basetable["neg", "well"])
basetable <- getposterior(newprior)
basetable["pos", "sick"] / (basetable["pos", "sick"] + basetable["pos", "well"])
```

This test compares to the answer in 5.1 (0.01943463) quite poorly... why?

### Exercise 5.4

Use BenGrid

```{r}
source("~/Documents/bookclub/DBDA2Eprograms/DBDA2E-utilities.R")  
source("~/Documents/bookclub/DBDA2Eprograms/BernGrid.R")
```

```{r}
Theta = seq( 0 , 1 , length=5 )  # Sparse teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,0),rep(1,1))      # Single flip with 1 head

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="None" , showHDI=FALSE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=11 )  # Sparse teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,0),rep(1,1))      # Single flip with 1 head

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="None" , showHDI=FALSE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1001 ) # Fine teeth for Theta.
pTheta = rep(1,length(Theta))      # Uniform (horizontal) shape for pTheta.
pTheta = pTheta/sum(pTheta)        # Make pTheta sum to 1.0
Data = c(rep(0,0),rep(1,1))        # Single flip with 1 head

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="None" , showHDI=FALSE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1001 ) # Fine teeth for Theta.
pTheta = rep(0,length(Theta))      # Only extremes are possible!
pTheta[2] = 1                      # Only extremes are possible!
pTheta[length(pTheta)-1] = 1       
pTheta = pTheta/sum(pTheta)        # Make pTheta sum to 1.0
Data = c(rep(0,0),rep(1,1))        # Single flip with 1 head

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="None" , showHDI=FALSE , showpD=FALSE )
```

```{r}



Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,3),rep(1,1))      # 25% heads, N=4

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
pTheta = pTheta^10               # Sharpen pTheta !
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,3),rep(1,1))      # 25% heads, N=4

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
pTheta = pTheta^0.1              # Flatten pTheta !
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,3),rep(1,1))      # 25% heads, N=4

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

```{r}


Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,30),rep(1,10))    # 25% heads, N=40

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
pTheta = pTheta^10               # Sharpen pTheta !
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,30),rep(1,10))    # 25% heads, N=40

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1001 )  # Fine teeth for Theta.
pTheta = pmin( Theta , 1-Theta ) # Triangular shape for pTheta.
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
pTheta = pTheta^0.1              # Flatten pTheta !
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,30),rep(1,10))    # 25% heads, N=40

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="Mode" , showHDI=TRUE , showpD=FALSE )
```

```{r}

Theta = seq( 0 , 1 , length=1000 )  # Fine teeth for Theta.
# Two triangular peaks on a small non-zero floor:
pTheta = c( rep(1,200),seq(1,100,length=50),seq(100,1,length=50),rep(1,200) , 
            rep(1,200),seq(1,100,length=50),seq(100,1,length=50),rep(1,200) )
pTheta = pTheta/sum(pTheta)      # Make pTheta sum to 1.0
Data = c(rep(0,13),rep(1,14)) 

posterior = BernGrid( Theta, pTheta , Data , plotType="Bars" , 
                      showCentTend="None" , showHDI=FALSE , showpD=FALSE )
```
