---
title: "bayes-ch7"
output: html_document
---

# Chapter 6

### Exercise 6.1-A

First let's clean up the code from `BernMetrop.R` from the book.

This will set up a few convenience functions:

```{r}
# Define the Bernoulli likelihood function, p(D|theta).
# The argument theta could be a vector, not just a scalar.
likelihood <- function(theta, data) {
  z <- sum(data)
  N <- length(data)
  pDataGivenTheta <- theta^z * (1-theta)^(N-z) # Bernoulli
  # The theta values passed into this function are generated at random,
  # and therefore might be inadvertently greater than 1 or less than 0.
  # The likelihood for theta > 1 or for theta < 0 is zero:
  pDataGivenTheta[ theta > 1 | theta < 0 ] <- 0
  pDataGivenTheta
}

# Define the prior density function. 
prior <- function(theta) {
  pTheta <- dbeta(theta, 1, 1)  # dbeta from stats package
  # The theta values passed into this function are generated at random,
  # and therefore might be inadvertently greater than 1 or less than 0.
  # The prior for theta > 1 or for theta < 0 is zero:
  pTheta[ theta > 1 | theta < 0 ] <- 0
  pTheta
}

# Define the relative probability of the target distribution, 
# as a function of vector theta. For our application, this
# target distribution is the unnormalized posterior distribution.
targetRelProb <- function(theta, data) {
  likelihood(theta, data) * prior(theta)
}

```

Set up a metropolis function (original code had this as stand alone)

```{r}

# Specify the data, to be used in the likelihood function (myData)
# Specify the length of the trajectory, i.e., the number of jumps to try (trajLength)
# Specify standard deviation of proposal distribution (proposalSD)
metrop <- function(myData, proposalSD=0.2, trajLength=50000) {
  # Initialize the vector that will store the results:
  trajectory <- rep(0, trajLength)
  # Specify where to start the trajectory:
  trajectory[1] <- 0.01 # arbitrary value
  # Specify the burn-in period:
  burnIn <- ceiling(0.0 * trajLength) # arbitrary number, less than trajLength
  ## Note: how is BurnIn not always zero?

  # Initialize accepted, rejected counters, just to monitor performance:
  nAccepted <- 0
  nRejected <- 0

  # Now generate the random walk. The 't' index is time or trial in the walk.
  for ( t in 1:(trajLength-1) ) {
    currentPosition = trajectory[t]
    # Use the proposal distribution to generate a proposed jump.
    proposedJump = rnorm( 1 , mean=0 , sd=proposalSD )
    # Compute the probability of accepting the proposed jump.
    probAccept = min( 1,
                      targetRelProb( currentPosition + proposedJump , myData )
                      / targetRelProb( currentPosition , myData ) )
    # Generate a random uniform value from the interval [0,1] to
    # decide whether or not to accept the proposed jump.
    if ( runif(1) < probAccept ) {
      # accept the proposed jump
      trajectory[ t+1 ] = currentPosition + proposedJump
      # increment the accepted counter, just to monitor performance
      if ( t > burnIn ) { nAccepted = nAccepted + 1 }
    } else {
      # reject the proposed jump, stay at current position
      trajectory[ t+1 ] = currentPosition
      # increment the rejected counter, just to monitor performance
      if ( t > burnIn ) { nRejected = nRejected + 1 }
    }
  }
  
  # Extract the post-burnIn portion of the trajectory.
  acceptedTraj = trajectory[ (burnIn+1) : length(trajectory) ]
  list(acceptedTraj=acceptedTraj, 
       proposalSD=proposalSD, 
       trajectory=trajectory, 
       nAccepted=nAccepted, 
       nRejected=nRejected, 
       trajLength=trajLength,
       burnIn=burnIn)
}
# End of Metropolis algorithm.
```

Now we can run the `metrop` function with any parameters

But first, let's fix up the plotting function and make it a function as well.

Also note that I changed the layout here and made it show the first 200 and last 200 as well.

```{r}
source("~/Documents/bookclub/DBDA2Eprograms/DBDA2E-utilities.R")

plot.metrop <- function(myData, proposalSD=0.2, trajLength=50000) {
  r1 <- metrop(myData, proposalSD, trajLength)
  # layout( matrix(1:3,nrow=3) )
  layout(matrix(c(1, 1, 2, 3, 2, 3), 3, 2, byrow = TRUE))
  par(mar=c(3,4,2,2),mgp=c(2,0.7,0))

  # Posterior histogram:
  paramInfo <- plotPost(r1$acceptedTraj, xlim=c(0,1), xlab=bquote(theta), 
                        cex.main=2.0,
                        main=bquote( 
                          list("Prpsl.SD"==.(r1$proposalSD),
                               "Eff.Sz."==.(round(effectiveSize(r1$acceptedTraj),1)))))
  # Trajectory, a.k.a. trace plot, end of chain:
  idxToPlot = (r1$trajLength-200):r1$trajLength
  plot( r1$trajectory[idxToPlot] , idxToPlot , main="End of Chain" ,
        xlab=bquote(theta) , xlim=c(0,1) , ylab="Step in Chain" ,
        type="o" , pch=20 , col="skyblue" , cex=0.75, cex.lab=1.5 )
  # Display proposal SD and acceptance ratio in the plot.
  text( 0.0 , r1$trajLength , adj=c(0.0,1.1) , cex=1.75 ,
        labels = bquote( frac(N[acc],N[pro]) == 
                           .(signif( r1$nAccepted/length(r1$acceptedTraj) , 3 ))))
  
  # Trajectory, a.k.a. trace plot, beginning of chain:
  idxToPlot = 1:200
  plot( r1$trajectory[idxToPlot] , idxToPlot , main="Beginning of Chain" ,
        xlab=bquote(theta) , xlim=c(0,1) , ylab="Step in Chain" ,
        type="o" , pch=20 , col="skyblue", cex=0.75, cex.lab=1.5 )
  # Indicate burn in limit (might not be visible if not in range):
  if ( r1$burnIn > 0 ) {
    abline(h=r1$burnIn,lty="dotted")
    text( 0.5 , r1$burnIn+1 , "Burn In" , adj=c(0.5,1.1) )
  }
  layout(matrix(1))
  invisible(r1)
}
```


```{r fig.height=7, fig.width=6}
set.seed(1)
r1 <- plot.metrop(c(rep(0,6),rep(1,14)), proposalSD=0.3, trajLength=50000)
```

```{r fig.height=7, fig.width=6}
set.seed(1)
r1 <- plot.metrop(c(rep(0,6),rep(1,14)), proposalSD=0.2, trajLength=50000)
```

```{r fig.height=7, fig.width=6}
set.seed(2)
r1 <- plot.metrop(c(rep(0,6),rep(1,14)), proposalSD=0.1, trajLength=50000)
```

```{r fig.height=7, fig.width=6}
set.seed(2)
r1 <- plot.metrop(c(rep(0,6),rep(1,14)), proposalSD=0.02, trajLength=10000)
```

Just for fun, I want to play with that "Effective Size" bit and see where it 
is maximized.

```{r cache=TRUE}
outplot <- do.call(rbind, lapply(seq(0.1, 0.3, by=0.005), function(proposalSD) {
  eff <- mean(sapply(seq(10), function(i) {
    effectiveSize(metrop(c(rep(0,6),rep(1,14)), proposalSD, trajLength=5000)$acceptedTraj)
  }))
  data.frame(sd=proposalSD, eff=round(eff, 1))
}))
suppressPackageStartupMessages(library(ggplot2))
ggplot(outplot, aes(sd, y=eff)) + geom_point() + geom_smooth(se=FALSE) + theme_bw()

```